---
title: "Regression Models Course Project"
author: "Rodrigo Falcão"
date: "`r Sys.Date()`"
output: pdf_document
---
# Exploratory Data Analysis
```{r setup, include=FALSE}
library(GGally)
library(ggplot2)
data(mtcars)
```

First we have to cast categorical variables to factor and check the box plot between "mpg" and "am".

``` {r}
factor_cols <- c("vs", "gear", "carb")
mtcars[factor_cols] <- lapply(mtcars[factor_cols], factor)
mtcars$am <- factor(mtcars$am, labels=c("Automatic","Manual"))

boxplot(mpg ~ am, data = mtcars, col = (c("red","blue")), ylab = "Miles Per Gallon", xlab = "Transmission Type")
```
From the plot is it possible to see that the automatic transmissions tends to be worse than manual transmission when it comes to fuel consumption (mpg). 
# Statistical Inference

``` {r}
t.test(mpg ~ am, mtcars)
```
From the T test is it possible to see that manual transmission is better when it comes to "mpg". But since the mean value itself does not represent a concrete proof that the manual transmission is better it is necessary to fit the data to a regression model. Let's start fitting just with the factor variable of transmission.

# Regression model
``` {r}
fit_am <- lm(mpg ~ factor(am), data=mtcars)
summary(fit_am)
```
From the summary is it possible to see that the P-value is less than 0.0003, so the hypothesis is not rejected. Although the P-value is low the R-squared value for this test is approximately 0.35, indicating that around one-third of the variation in "mpg" can be explained by transmission type alone. As a next step, let's conduct an Analysis of Variance for the dataset.
``` {r}
fit_full <- lm(mpg ~ ., data=mtcars)
summary(fit_full)
```

The full model has an R-squared value of 0.8867, as expected. However, based on the summary output, none of the coefficients are statistically significant at the 0.05 level.

Removing variables that are correlated with transmission type could result in biased coefficients, while adding unnecessary regressors can increase the model's variance. To determine the appropriate variables to include in our final model, we will use the step function in R.

``` {r}
fit_optimal <- step(fit_full, direction = "backward", trace = FALSE)
summary(fit_optimal)
```
The step algorithm generated a model that includes the variables "wt", "qsec", and "am." All three variables have a significant impact at the 0.05 level, and the model accounts for around 85% of the variation.

Based on the coefficients of this model, we can infer that manual cars, on average, achieve 2.9 more miles per gallon than automatic cars, assuming "weight" and "qsec" remain constant.

The base graphics diagnostic plots demonstrate that there is no correlation between the residuals and the fitted values. Additionally, the quantile-quantile plot suggests that the residuals follow a normal distribution.

```{r}
par(mfrow = c(2,2))
plot(fit_optimal)
```

# Appendix

```{r, message=FALSE}
ggpairs(mtcars, mapping = aes(colour = am))
```